---
title: "vignette"
author: "Nijiati Abulizi, Kulaphong Jitareerat, Craig Adlam"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# YelpViz API Wrapper   

### Getting Started

Installation of the package:

```{r}
library(yelpviz)
```

Install required libraries:

```{r}
library(httr)
library(jsonlite)
library(dplyr)
library(ggplot2)
library(plotly)
library(stringr)
```

### analyze_cities()

To analyze the ratings for different cities based on a single business sector within an interactive density plot, call the analyze_cities function. This retrieves the best matched businesses for that business sector and can accept multiple cities for comparison. For quick analysis, simply call `analyze_cities()` and enter required information when prompted.

**Prompts**\
*"Please enter your Yelp API key: "*\
*"Please enter the cities (separated by commas if multiple): "*\
*"Please enter the category name: "*

***The dataframe, parameter values, and plot are returned.***

```{r}
analyze_cities()
```

***--- discuss output here ---***

Or for more in-depth analysis, the user can set or enter the parameters prior and assign to a variable to be called:\
\
**Required parameters**

| Parameter  | Description          | Default |
|------------|----------------------|---------|
| `api_key`  | Yelp API key         | NULL    |
| `cities`   | List of cities       | NULL    |
| `category` | Business sector      | NULL    |
| `limit`    | Number of businesses | 20      |

```{r}
# Enter the Yelp API key and specify multiple cities in a vector
api_key <- <<<ENTER API KEY HERE>>>
cities <- c('Kelowna', 'Red Deer', 'Toronto')

# Assign to a variable for further analysis
result <- analyze_cities(api_key, cities, 'food', 33)
#result$combined_df # uncomment to see the dataframe
#result$parameters # uncomment to see the parameters
#result$plot # uncomment to see the plot
```

Now the user has the option to create other visualizations, retrieve summary statistics, or view the ratings, review counts, and prices of each business within the business sector.

For example, the mean rating of each city can accessed:

```{r}
# Convert 'rating' column to numeric
result$combined_df$rating <- as.numeric(result$combined_df$rating)

# Calculate mean rating for City = 'Red Deer'
mean_rating_red_deer <- mean(result$combined_df$rating[result$combined_df$City == "Red Deer"], na.rm = TRUE)
mean_rating_red_deer
```

***--- discuss output here ---***

### analyze_business_sectors()

To analyze the ratings for different business sectors based on a single city by displaying a stacked-density plot, call the analyze_business_sectors function. This retrieves the best matched businesses for each business sector and can accept a single city for comparison. For quick analysis, simply call `analyze_business_sectors()` and enter required information when prompted.

**Prompts**\
*"Please enter your Yelp API key: "\
"Please enter the city name: "\
"Please enter the categories (separated by commas if multiple): "*

***The dataframe, parameter values, and plot are returned.***

```{r}
analyze_business_sectors()
```

***--- discuss output here ---***

Or for more in-depth analysis, the user can set or enter the parameters prior and assign to a variable to be called:\
\
**Required parameters**

| Parameter    | Description              | Default |
|--------------|--------------------------|---------|
| `api_key`    | Yelp API key             | NULL    |
| `city`       | Single city              | NULL    |
| `categories` | List of business sectors | NULL    |
| `limit`      | Number of businesses     | 20      |

```{r}
# Enter the Yelp API key and specify multiple categories in a vector
api_key <- <<<ENTER API KEY HERE>>>
categories <- c('food', 'gyms', 'golf')

# Assign to a variable for further analysis
result <- analyze_cities(api_key, 'Kelowna', categories, 33)
#result$combined_df # uncomment to see the dataframe
#result$parameters # uncomment to see the parameters
#result$plot # uncomment to see the plot
```

Now the user has the option to create other visualizations, retrieve summary statistics, or view the ratings, review counts, and prices of each business sector within the city.

For example, the rating and review count summary statistics of each business sector can accessed:

```{r}
summary(result$combined_df)
```

***--- discuss output here ---***

***--- consider discussing other visualization output here ---***
